---
title: "  What I read in 2020: My ratings versus Goodreads"
description: |
  How well do my book ratings compare to the Goodreads community ratings? 
author:
  - name: Gary Lynch
    url: https://olytico.com/
    affiliation: Olytico
    affiliation_url: https://olytico.com/
date: 01-01-2021
output:
  distill::distill_article:
    self_contained: false
---
## Introduction

2020 ended up being a uniquely good year for reading (at least for me). A series of lockdown months helped to reduce the number of distractions I was exposed to. There wasn't much rugby going on, cinemas were closed and pubs rapidly became a distant memory. 

As I took a look back at what I read in 2020, it struck me that it might be interesting to see whether the wider reading public had the similar views on what I've read. The best proxy I could find for the wider reader consensus on a book was the average review scores from Goodreads.com. I plotted my personal ratings against the

Each point on the graph represents a book, with the y-axis showing my rating and the x-axis gives the Goodreads rating. Any point that sits on the line represents a book where my rating and the Goodreads rating matched. You can hover over any point to see it's title and author.

```{r rbokeh1, echo=FALSE}

if(!require("rbokeh")) install.packages("rbokeh",repos = "http://cran.us.r-project.org")
library(rbokeh)

if(!require("kableExtra")) install.packages("kableExtra",repos = "http://cran.us.r-project.org")
library("kableExtra")

if(!require("ggplot2")) install.packages("ggplot2",repos = "http://cran.us.r-project.org")
library("ggplot2")

BookData <- read.csv("C:/Users/Gary/Documents/BookData.csv")
```

```{r cars}

graph11 <- figure(legend_location = "top_left", plot_width = 1000, plot_height = 1000, tools = "Null", xlab = "Good Reads Rating", ylab = "My Rating") %>%
  ly_points(Good.Reads.Rating, My.Rating, data = BookData, color = Type,
##Set the colour based on the type variable and the glyph/symbol to the treatment variable
    hover = list(Title, Author)) %>%
 ly_abline(0, 1)
graph11
```

As you can see, only 2 of 40 books had a matched rating. That said, most of the ratings were fairly close, save for a few notable exceptions. Those were the books I was most interested in - where I had a significantly different opinion from the group consensus. Below I've pulled out the books with the greatest difference from the Goodreads rating. They're split into books I significantly preferred and books that I was significantly less keen on. 

```{r cardds}
BookData <- read.csv("C:/Users/Gary/Documents/BookData.csv")

BookData2 <- BookData[-c(6)]

BookDataminus <- BookData2[with(BookData,order(-Difference)),]

BookDataplot <- BookData2[with(BookData,order(Difference)),]

kbl(BookDataplot[1:5,])%>%  
  kable_styling(c("striped", "hover","responsive"))
```


```{r carddsss}

kbl(BookDataminus[1:5,])%>%  
  kable_styling(c("striped", "hover","responsive"))
```

The Memory Police: This was hands down the best book I read in 2020. Definitely not for everyone, which I think explains the big difference in ratings. If you enjoy allegorical fiction or existential storytelling, I can recommend.

If Cats Disappeared from the World: This is one of those books that sneaks up on you. It begins in a lighthearted manner and slowly reveals its emotional core as you go. I was genuinely surprised its rating wasn't higher.

Normal People: So good that its now a major television series. The book hit very close to home for me, not least because its set in the university I attended. Worth reading for the culture impact alone - or you could just watch the Tv show.

The Color Purple: A classic for a reason. This moved me more than anything else this year and feels very relevant to current day issues for a novel from the 1980's.

Frankenstein: A classic with a much lower Goodreads rating than I expected. A landmark novel that is surpisingly readable for something written over 200 years ago.

```{r car98idds}

kbl(BookDataplot[1:5,])%>%  
  kable_styling(c("striped", "hover","responsive"))
```

Little Fires Everywhere: Now a major television series. A novel that I may have had too high expectations for. Many loved this but for me


```{r casadsars}
crackd <- c(table(BookData$Month.Read))

crackd <- c(crackd,0)

names(crackd)[12] <- "October"

crackd2 <- crackd[c(5, 4, 8, 1, 9, 7, 6, 2, 11,12,10,3)]

names(crackd2) <- c("Jan","Feb","Mar","Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

barplot(crackd2, cex.names=0.9, col = "lightslateblue")

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating blogs with Distill at:
# https://rstudio.github.io/distill/blog.html

```



