---
title: "What I read in 2020: My ratings versus Goodreads"
description: |
  How well do my book ratings line up to the Goodreads community ratings? 
preview: https://raw.githubusercontent.com/Gary-Lynch/gary-lynch.github.io/main/posts/2021-01-16-what-i-read-in-2020-my-ratings-versus-goodreads/what-i-read-in-2020-my-ratings-versus-goodreads_files/figure-html5/rbokeh2-1.png
author:
  - name: Gary Lynch
    url: https://gary-lynch.github.io/
date: 01-16-2021
output:
  distill::distill_article:
    self_contained: false
---

## Introduction

2020 ended up being a uniquely good year for reading (at least for me). Multiple lockdown left me with lots of reading time and for the first time I kept track of what I read during the year. 

I thought it would be interesting to share some of what I tracked by answering a couple of questions. Mainly, I want **to see how my ratings compare to the average reader** (for this, I used Goodreads average reviews). I don't use the site myself, but its a great proxy for the general consensus on a book. 

The questions I'd like to answer are as follows:

- How did my ratings compare with Goodreads average ratings?
- What books did I like more than Goodreads readers?
- What books did Goodreads readers like more than me?

```{r rbokeh1, echo=FALSE}

if(!require("distill")) install.packages("distill",repos = "http://cran.us.r-project.org")
library(distill)

if(!require("rbokeh")) install.packages("rbokeh",repos = "http://cran.us.r-project.org")
library(rbokeh)

if(!require("kableExtra")) install.packages("kableExtra",repos = "http://cran.us.r-project.org")
library("kableExtra")

if(!require("ggplot2")) install.packages("ggplot2",repos = "http://cran.us.r-project.org")
library("ggplot2")

BookData <- read.csv("C:/Users/Gary/Documents/BookData.csv")
```

## How do my ratings stack up against Goodreads?
```{r rbokeh2, include=FALSE}
plot(BookData$Good.Reads.Rating,BookData$My.Rating, col = "cyan3", pch = 16,xlab = "Good Reads Community Rating", ylab = "My Rating")
abline(0,1)
```
The below graph represents the 40 books I read in 2020. Each point relates to one book, with the horizontal axis showing the Goodreads average rating and the vertical axis showing my rating. I've added a line to highlight the points where my rating matches the Goodreads rating (Note: You can hover over any point to see the book it relates to).

\ \

```{r cars, echo=FALSE}

graph11 <- figure(legend_location = "top_left", plot_width = 1000, plot_height = 1000, tools = "Null", xlab = "Good Reads Rating", ylab = "My Rating") %>%
  ly_points(Good.Reads.Rating, My.Rating, data = BookData, color = Type,
##Set the colour based on the type variable and the glyph/symbol to the treatment variable
    hover = list(Title, Author)) %>%
 ly_abline(0, 1)
graph11
```

As you can see, only 2 of 40 books had a matched rating. That said, most of the ratings were fairly close, with the majority within 0.5 of the other rating. There were a few notable exceptions, which I'll go into more detail on below.

I was surprised to see that nearly every non-fiction book I read was rated higher on Goodreads. I have a feeling this has more to do with Goodreads ratings than to do with me (did you notice that the majority of +4 star Goodreads reviewed books are non-fiction?). Maybe I just exclusively read great non-fiction?

In the next section, we'll look at the books with the greatest difference in ratings - starting with books I liked more than Goodreads.

## Books I liked more than Goodreads

```{r cardds, echo=FALSE}
BookData <- read.csv("C:/Users/Gary/Documents/BookData.csv")

BookDataminus <- BookData[with(BookData,order(-Difference)),]

BookDataplot <- BookData[with(BookData,order(Difference)),]

kbl(BookDataminus[1:5,1:5,])%>%  
  kable_styling(c("striped", "hover","responsive")) %>%  
  column_spec(1, bold = T)
```

**The Memory Police:** This was hands down the best book I read in 2020, so I expected this to exceed the Goodreads rating. Definitely not a book for everyone, but a good pick if you enjoy allegorical fiction.

**If Cats Disappeared from the World:** A short read that packs an emotional punch. It starts quite lighthearted and slowly reveals itself to you. I was quite surprised that Goodreads readers didn't like this more.

**Normal People:** The best-selling novel that is now a major television series. This hit very close to home for me, not least because its set in the university I went to. Worth reading for the culture impact alone or as preparation for watching the TV show.

**The Color Purple:** Alice Walker's Pulitzer prize winning novel is a classic for a reason. This moved me more than anything else I read this this year. I found it remarkably relevant to current day issues for a novel from the 1980's.

**Frankenstein:** Mary Shelley's classic had a much lower Goodreads rating than I expected. A landmark novel that is very readable for something written over 200 years ago.

## Books preferred by Goodreads readers

```{r car98idds, echo=FALSE}

kbl(BookDataplot[1:5,1:5,])%>%  
  kable_styling(c("striped", "hover","responsive")) %>%  
  remove_column(1) %>%  
  column_spec(1, bold = T)
```

**Little Fires Everywhere:** Another best-selling novel that is now a major television series. I think I may have had too high expectations for this. I didn't hate this but the story just went in a direction story-wise that I didn't enjoy.

**Freedom:** I didn't realise when I bought this that it's actual a collection of excerpts from other Atwood novels. Unfortunately, I had read most of these novels before. Not an awful book, but I'd recommend just reading her actual novels instead.

**The Cuckoo's Calling:** This is the first in a series of murder mystery novels from J.K. Rowling. I found it fine as a read, if a little long. Obviously Goodreads readers enjoyed it more than I did.

**When All is Said:** Anne Griffin does a great job capturing life in Ireland here. I enjoyed reading this, but it was closer to a 3 than a 4 star for me.

**Sputnik Sweetheart:** I am a big Murakami fan. This was not my favorite of his, but it does have some beautiful description. If you're interested in trying Murakami, I would probably start with something else (Norwegian Wood, Kafka on the Shore).

## Final thoughts

This was a fun little experiment in plotting my reading for 2020. As a bonus, I've included the number of books I read each month below. You can probably guess when college started based on how my reading fell off a cliff!

```{r casadsars, echo=FALSE}
crackd <- c(table(BookData$Month.Read))

crackd <- c(crackd,0)

names(crackd)[12] <- "October"

crackd2 <- crackd[c(5, 4, 8, 1, 9, 7, 6, 2, 11,12,10,3)]

names(crackd2) <- c("Jan","Feb","Mar","Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

barplot(crackd2, cex.names=0.9, col = "deepskyblue3")

```
If you're interested in doing something similar to this, all the graphs were created in R Markdown. The first graph utilized the Rbokeh package and the tables use the Kable extra package. This blog was compiled using the distill package and is hosted on Github pages.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating blogs with Distill at:
# https://rstudio.github.io/distill/blog.html

```

Distill is a publication format for scientific and technical writing, native to the web.

Learn more about using Distill at <https://rstudio.github.io/distill>.


