---
title: "Game of the Year nominees - What got r/Gaming talking"
description: |
  Which of the Game of the Year nominees drove the most conversation on r/Gaming?
author:
  - name: Gary Lynch
    url: https://gary-lynch.github.io/
date: 01-27-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r rbokeh1, echo=FALSE, out.width="940px"}

if(!require("tidyverse")) install.packages("tidyverse",repos = "http://cran.us.r-project.org")
library(tidyverse)

if(!require("rbokeh")) install.packages("rbokeh",repos = "http://cran.us.r-project.org")
library(rbokeh)

if(!require("png")) install.packages("png",repos = "http://cran.us.r-project.org")
library("png")

if(!require("grid")) install.packages("grid",repos = "http://cran.us.r-project.org")
library("grid")

if(!require("ggplot2")) install.packages("ggplot2",repos = "http://cran.us.r-project.org")
library("ggplot2")

#> Linking to ImageMagick 6.9.9.39
#> Enabled features: cairo, fontconfig, freetype, lcms, pango, rsvg, webp
#> Disabled features: fftw, ghostscript, x11

  img <- readPNG("C:/Users/Gary/Pictures/Zagreus.png")
 g <- rasterGrob(img, interpolate=TRUE)

   img1 <- readPNG("C:/Users/Gary/Pictures/Cloud.png")
 gd <- rasterGrob(img1, interpolate=TRUE)
 
   img2 <- readPNG("C:/Users/Gary/Pictures/TLOU.png")
 gh <- rasterGrob(img2, interpolate=TRUE)
 
    img3 <- readPNG("C:/Users/Gary/Pictures/Doom.png")
 gr <- rasterGrob(img3, interpolate=TRUE)
 
     img4 <- readPNG("C:/Users/Gary/Pictures/GhostT.png")
 gf <- rasterGrob(img4, interpolate=TRUE)
 
      img5 <- readPNG("C:/Users/Gary/Pictures/ACNH.png")
 gs <- rasterGrob(img5, interpolate=TRUE)
 
      img6 <- readPNG("C:/Users/Gary/Pictures/Logo Files/Screen/LOGO_RGB_Transparent.png")
 grt <- rasterGrob(img6, interpolate=TRUE)
 
 yy <- c(15,12,22,34,55)
 xx <- c("The Last of Us", "Hades", "Fury", "Game 4", "Cyberpunk 2077")
 
data <- data.frame(
  name=c("Hades", "FF 7 Remake", "Doom: Eternal", "Ghost of Tsu.","TLOU: 2", "AC: NH"),  
  value=c(3913,24060,25113,26925,33113,36205)
  )
 
ggplot(data, aes(x=reorder(name,value), y=value)) + 
      annotation_custom(g, xmin=0.5, xmax=1.5, ymin=5000, ymax=19000) +
      annotation_custom(gd, xmin=1.5, xmax=2.5, ymin=25000, ymax=37000) +
  annotation_custom(gr, xmin=2.5, xmax=3.5, ymin=26000, ymax=37000) +
  annotation_custom(gf, xmin=3.5, xmax=4.5, ymin=28000, ymax=40000) +
     annotation_custom(gh, xmin=4.5, xmax=5.5, ymin=34000, ymax=42000) +
    annotation_custom(gs, xmin=5.5, xmax=6.5, ymin=36000, ymax=50000) +
     annotation_custom(grt, xmin=0.3, xmax=1.3, ymin=40000, ymax=50000) +
  geom_bar(stat = "identity", width = 0.6, fill="royalblue4") +
  ylab("Mentions on r/Gaming") +
  ggtitle("Game of the Year nominees: which titles got r/Gaming talking?") +
  ylim(0,45000) + 
 theme(axis.text.x = element_text(angle = 0, size=9, face="bold"),
  axis.title.x = element_blank(),
  axis.text.y = element_text(size=10, face="bold"),
  plot.title = element_text(hjust = 0.5,size=14, face="bold"),
  axis.title.y = element_text(size=12, face="bold")
       )
ggsave("GamAwardsHQ2.png", width = 31.75, height = 16.5, units = "cm")

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
## Introduction

The Game Awards are the gaming community's annual celebration of the best titles from the year. And 2020 had no shortage of great games to recognize. 

There were AAA heavy hitters, indie surprise hits and quarantine party games that kept us playing all year. In fact, with so many of us stuck at home for most of the year, games may have played a bigger role in entertainment than ever before.

2020 was also no stranger to controversy in the gaming world. Whether it was disastrous launches or surprise lockdown favourites, there was no shortage of gaming news to get the internet talking. And a lot of gamers did that talking on the subretddit, r/Gaming.

Given there was so much to talk about during the year, **did the Game Awards "Game of the Year" nominees make a splash on r/Gaming?** Or were they drowned out by bigger gaming stories? We've taken a look at which of the nominated games got redditors talking, as well as some of the other biggest titles from the year.

## #6 Hades (3,913 mentions)

<center>

<iframe width="560" height="315" src="https://www.youtube.com/embed/Bz8l935Bv0Y" frameborder="0" allowfullscreen align=”middle” margin=”auto” ></iframe>

</center>

Independent developer Supergiant games' have some experience releasing critically acclaimed titles, and their most recent outing is no slouch. Hades is  a rogue-like dungeon crawler set in the world of Greek gods and myth.You play as Zagreus, the rebellious son of Hades, on a mission to escape the Underworld. 
The witty dialogue, fast-paced combat and... appealing characters made this a favourite among critics at the end of the year. r/Gaming users shared the enthusiasm of critics, with the conversation around the game being overwhelmingly positive. While this was the smallest discussion for any of the game awards nominees analysed, there were some key posts that drove the conversation:

- One gamer's DIY Hades -themed arcade build [(Link)](https://www.reddit.com/r/gaming/comments/knwdrg/finished_my_hades_arcade_build_what_do_you_guys/)
- Praise for Clancy Brown's performance in Hades and other titles [(Link)](https://www.reddit.com/r/gaming/comments/jhtsfp/amazing_how_clancy_brown_performed_all_these/)
- A cartoon post entitled "Why I recommend Hades" [(Link)](https://www.reddit.com/r/gaming/comments/jwn2tl/why_i_recommend_hades/)

## #5 Final Fantasy 7 Remake (24,060 mentions)

<center>

<iframe width="560" height="315" src="https://www.youtube.com/embed/eHdaWMP7D2s" frameborder="0" allowfullscreen></iframe>

</center>

Since it was first announced in 2015, there has been a lot of hype for the Final Fantasy 7 remake. Based on the reception it received on release and over 24,000 mentions on r/Gaming, it seems this title delivered on that hype.

Early in the year, the game's conversation was dominated by its demo [(Link)](https://www.reddit.com/r/gaming/comments/fd2lp2/my_experience_with_ff7_remake_demo/fjevmj5/), and then moved to focus on disappointment at the game's postponement. The April release date saw this conversation hit it's peak, with a post featuring a fan's mother recreating an NPC's look driving mentions [(Link)](https://www.reddit.com/r/gaming/comments/g0m6g8/sent_a_picture_of_a_final_fantasy_7_character_to/).

## #4 Doom Eternal (25,113 mentions)

<center>
<iframe width="560" height="315" src="https://www.youtube.com/embed/_J2YcaVqjCQ" frameborder="0" allowfullscreen></iframe>

</center>

Following up from the surprise hit that was Doom 2016 was never going to be easy, but ID software seem to have done it. The Doom Slayer returns in the 5th installment to the Doom series, for one of the biggest games of 2020.

Releasing as quarantines took effect around the world, the game received strong praise from gamers. Humorous content drove a decent chunk of conversation at release [(Link)](https://www.reddit.com/r/gaming/comments/flvbl9/this_is_the_top_review_for_the_new_doom_eternal/fl7mmma/). Another key driver of mentions was a post highlighting the effect playing Doom Eternal had on one players heart rate [(Link)](https://www.reddit.com/r/gaming/comments/fniz7x/so_i_played_some_doom_eternal_this_weekend_and_my/).

## #3 Ghosts of Tsushima (26,952 mentions)

<center>
<iframe width="560" height="315" src="https://www.youtube.com/embed/MamonBeCaAQ" frameborder="0" allowfullscreen></iframe>

</center>

Samurai games have seen a surge in recent years, and Ghost of Tsushima is the most recent action-adventure game to opt for the setting. Sucker Punch productions, known for their work on the Infamous series, have taken that formula and crafted one of the fastest selling original titles in recent memory.

Upon release, there was a conversation on r/Gaming focused on the impressive visuals of the game [(Link)](https://www.reddit.com/r/gaming/comments/jq2z0j/just_started_playing_ghost_of_tsushima_it_is_one/). Discussions of historical accuracy also featured  in the weeks following the launch [(Link)](https://www.reddit.com/r/gaming/comments/htgi4d/weird_western_gamers_vs_cool_eastern_gamers_about/). 

## #2 The Last of Us: Part 2 (33,113 mentions)

<center>
<iframe width="560" height="315" src="https://www.youtube.com/embed/II5UsqP2JAk" frameborder="0" allowfullscreen></iframe>

</center>

The sequel that will be remembered as one of the most controversial titles from 2020. Or maybe ever.

The Last of Us: Part 2 was a title that divided gamers and critics alike, and arguments about the game's quality drove mentions on r/Gaming. Redditors opinions ranged from game of the year to worst game ever made, and covered everything in between. 

Ultimately, the title was always going to be compared to its predecessor and we saw that conversation play out on the subreddit [(Link)](https://www.reddit.com/r/gaming/comments/hc419h/holy_shit/). 

Despite the divisive discourse, it was a much happier game that topped our list as the most discussed nominee on r/Gaming.

## #1 Animal Crossing: New Horizons (36,305 mentions)

<center>
<iframe width="560" height="315" src="https://www.youtube.com/embed/u9TY741PSh8" frameborder="0" allowfullscreen></iframe>

</center>

The social game that dominated quarantines and lockdowns across the globe - Animal Crossing: New Horizons was the most discussed Game of the Year nominee on r/Gaming. 

For many, playing Animal Crossing represented the best way to interact with friends at a time when doing that in person wasn't possible. Whether it was by happenstance or design, the title launched in a moment when social gaming was in high demand. And that certainly contributed to its success.


## Bonus Round - Cyberpunk 2077, Among Us, Fall Guys, Half-Life: Alyx

You really can't talk about the big games of 2020 without mentioning Cyberpunk 2077. Despite launching just 21 days before the year ended, there were over **210,000 mentions** of the game on r/Gaming. If you want to get talked about, a bug-filled launch is one way to do it.

Among Us was another social game that rose to prominence during 2020. When everyone found themselves stuck inside, private games made for a great replacement to zoom calls (remember those?). The game was **mentioned almost 45,000 times** on the subreddit, hitting a peak in April.

How many of the big games this year were lockdown favourites? It may have had something to do with all the time we all spent... in lockdowns. In any case, Fall Guys launched in August and for a while, it felt like the only game anyone could talk about. And then the gaming world moved on. This was reflected on r/Gaming, where it was *mentioned over 19,000 times*, mainly during August and September.

The last bonus game was released in March of 2020, and represented the end of waiting for many gamers. Half-Life:Alyx may not have been the much anticipated Half-Life 2, but it certainly made an impression on gamers. The VR title drove approx. **13,000 mentions during 2020**, with gifs and videos displaying the game's physics engine dominating the conversation [(Link)](https://www.reddit.com/r/gaming/comments/fnv8kh/as_an_extremely_immature_man_this_sold_the_half/). 

-----
The data for this blog post was sourced from Olytico - you can find out more at [here](https://olytico.com).

If you have any questions or would like to know more about any aspect of this blog post, you can reach me on my [Twitter account](https://twitter.com/GaryLynch94). 


